---
title: "*Absalom, Absalom!* and the Use of Racial Words"
author: "Johannes Burgers"
output:
  html_document:
    theme: yeti
    df_print: paged
bibliography: "bibliography.bib"
---


```{r setup, include=FALSE, message = FALSE, warning=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, error=FALSE, tidy='styler')
```

#### Load Libraries
```{r load_packages}
library(tidyverse)
library(tidytext)
library(stringi)
library(scales)
library(stringr)
library(plotly)
library(textstem)
library(shiny)
library(lubridate)
library(htmlTable)
```

#### Load texts
```{r load_text}
#Read in the text
absalom_df <- as.data.frame(read_file("absalom_cleaned_to_match_4_30_rev4.txt"))
colnames(absalom_df) <- "text" 

absalom_df <- absalom_df %>% 
mutate(text = gsub("[‘’]", "'", text))
```

```{r _load_event_sentences}
absalom_event_sentences <- read_csv("absalom_event_sentences_clean_2021_5_10.csv")

#This file was created by recomposing the text based on the "First Words" column and matching the text segments with each event using a copy of the text. As this process is quite involved it has been branched out into a separate file: absalom_sentiment_text_processing. Full details on the process can be viewed there.

```

```{r absalom_event_word_count}
event_length <- absalom_event_sentences %>%
  mutate(event_words = str_count(event_sentence,  "\\S+")) %>%
  select(EventID, begin_word, end_word, event_sentence, event_words)
```


```{r read_all_absalom_events}
absalom_all_events <- read_csv("absalom_all_events_2021_5_11.csv")
```

```{r clean_up_absalom_events}
#The full table is long and hard to work with. 
absalom_events_short <- absalom_all_events %>%
  select(3:11, 13, 21:27, 50, 52:54) %>% 
  mutate(narrator = str_extract(Summary, "(?<=(SOURCE)).+[^]</p>]")) %>% 
  mutate(narrator = str_replace_all(narrator,":","")) %>% 
  left_join(event_length, by="EventID")
```

```{r read_all_DY_events}
dy_events <- read_csv("full_database_2021_5_11.csv")


```

```{r clean_up_all_events}
dy_events_short <- dy_events %>% 
                    select(1:11, 21:27, 50, 52:54)


```


```{r load_race_words}
#Read in the race words
race_words <- read_csv("race_words.csv")
```

#### Load Chart Vars

```{r load_color_scheme, echo=FALSE}
#Page styling

#This is where you create your own custom color palette for the traces.
faulkner_colorway = c("#132C53","#F27A18","#ae0700","#79b473","#38726c","#76bed0","#6b2d5c","#448b2d","#e6d812")

#This controls the background color for the entire chart. Probably best left white.
faulkner_paperbackground = c('rgba(255,255,255,0)')

#This controls the background for the plot. Probably best left white.
faulkner_plotcolor = c('rgba(255,255,255,.3)')

#Margin 

m <- list(l = 50, r = 50, b = 50, t = 50, pad = 4)

#Caption Style

fig_caption <- "font-family: 'Open Sans','Helvetica Neue',Helvetica,Arial,sans-serif; font-weight: normal; font-size:90%"

plot_font <- list(
  family = "'Open Sans','Helvetica Neue',Helvetica,Arial,sans-serif",
  size = 14,
  color = '#363636')



```



## Introduction


### Overview

The following piece uses several techniques available to standard CL analysis, alongside a more complex analysis exclusively available to practitioners who have access to the [*Digital Yoknapatawpha*](http://faulkner.iath.virginia.edu/) data set. These different techniques have been split into their own sections. 

- **Part 1: Characters**  
          
- **Part 2: Locations**  
          
- **Part 3: Events**  

- **Part 4: Text**  





  The *DY* database provides insight into what characters are where and at what time. Using this database, it is possible to reconstruct the language surrounding people, places, and times.^[All of the data was generated in the R programming language using the `tidyverse` suite of packages for the calculations and the `plotly` library for the graphics. The full repository is available at https://github.com/joostburgers/absalom_sentiment_analysis Due to copyright issues the repository does not include the *Absalom, Absalom!* text file used for data analysis. The text file used for text analysis was the 2011 Vintage Edition [@faulkner2011], which corresponds with the 1990 edition we used for *DY*]


## Part 1: Characters

One of the common descriptions of *Absalom, Absalom!* is that it is not a whodunit, but rather a whydoneit. The reader is afforded the basic facts of the case early on in the narrative: Henry Sutpen shoots his sister's fiance, Charles Bon, on the day of their wedding. The remainder of the narrative is dedicated to unspooling the various lines of inquiry into why Henry was motivated to shoot Charles Bon. The answer is as integral to the Sutpen family story as it is a synecdoche of the larger history of slavery in America and the way enslavement and its concomitant legal apartheid regime structured social relations. In many respects then, *Absalom, Absalom!* is, at bottom, a character study. Faulkner sketches out a few basic details about the Sutpen family, and leaves it to the various narrators to infer cause and effect. 

To understand all of these various social relations, it is useful to think through not only the individual characters, but the larger demography of the text. Digital Yoknapatawpha is particularly useful in this regard. Not only is it possible to pull up the biographies of each individual character, it is also possible to "slice" and "drill-down" into the data to answer to render a more detailed portrait of the various characters in the text in relation to one another. 


```{r general_overview}
unique_characters <- absalom_events_short %>%
  distinct(CharacterName)


peripheral_characters <- absalom_events_short %>%
  distinct(CharacterName, .keep_all = TRUE) %>%
  count(Rank) %>%
  mutate(percent = n / sum(n))


Judith_events <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  filter(CharacterName == "Judith Sutpen")

Judith_unique_all <- absalom_events_short %>%
  semi_join(Judith_events, by = "EventID") %>%
  filter(PresentMentioned == "Present") %>%
  distinct(CharacterName, .keep_all = TRUE) %>%
  filter(CharacterName != "Judith Sutpen")

Judith_unique <- Judith_unique_all %>%
  filter(Rank %in% c("Peripheral", "Minor")) %>%
  distinct(CharacterName)


```

Although much of the text is heavily focused on the Sutpen family, its word is actually peopled by `r nrow(unique_characters)` unique characters. The overwhelming majority of these (`r percent(peripheral_characters$percent[2]+peripheral_characters$percent[3])`) are peripheral or minor characters. While the abundance of minor characters in *Absalom, Absalom!* comports with novels more generally, it is important reminder that while the Sutpen's may be secluded they are not isolated from the larger world. Indeed, Judith is probably the most sheltered of the Sutpen's, she is seen very infrequently in town, and beyond this 12 mile journey never strays from the home. Still, over the course of the novel she appears in events with `r nrow(Judith_unique_all)` characters, `r nrow(Judith_unique)` of whom are minor or peripheral characters, such as Colonel Sartoris, Theophilus McCaslin, and the unnamed Men in Sartoris and Sutpen's Regiment. Mr. Compson's description of her as a "Country Girl Dirt Farmer" is therefore not entirely accurate. Beyond this rather trivial discrepancy, it speaks to the fact that Judith is not immune from social opprobrium, even if she is the daughter of the area's wealthiest planters and lives at some remove from town.

The source of Judith's social opprobrium is, of course, her potential marriage to Charles Bon, which Thomas fears and Henry prevents. To understand the unique source of anxiety surrounding it is important to underscore some of the basic demographics of the real Lafayette and the fictional Yoknapatawpha. In his thorough historical account of Lafayette county, Don Harrison Doyle shows that the ratio of Black (40%) and White (60%) identified people in the county was relatively stable throughout the counties history. In his own maps of Yoknapatawpha Faulkner actually flips this ratio and claims that there are 6,298 Whites and 9313 "Negroes." 

```{r black_white_ratios}

all_characters <- dy_events_short %>%
  distinct(CharacterName) %>%
  nrow()

black_white_all <- dy_events_short %>%
  # filter(SourceTextCode!="AA") %>%
  distinct(CharacterName, .keep_all = TRUE) %>%
  count(Race) %>%
  mutate(percent = n / sum(n))

black_white_pre <- dy_events_short %>%
  #filter(SourceTextCode!="AA") %>%
  filter(Era %in% c("Antebellum (1831-1860)", "Pre-Removal (-1830)")) %>%
  distinct(CharacterName, .keep_all = TRUE) %>%
  count(Race) %>%
  mutate(percent = n / sum(n))
```

```{r count_char}
char_count_AA <- absalom_events_short %>%
  distinct(CharacterName, Race) %>%
  count(Race)  %>%
  mutate(percent = n / sum(n))
              
```


He does not mirror this in his fiction though. If we look at all `r comma(all_characters)` characters, there are significantly more White characters, `r comma(black_white_all$n[10])` (`r percent(black_white_all$percent[10])`) than there are Black characters `r comma(black_white_all$n[2])` (`r percent(black_white_all$percent[2])`). In the context of *Absalom, Absalom!*, this ratio changes a bit for those events that pre-date the Civil War. Here there are relatively more Native American (`r percent(black_white_pre$percent[3])`) and Black (`r percent(black_white_pre$percent[2])`) characters, and therefore proportionally fewer White characters (`r percent(black_white_pre$percent[9])`). Thus, even though much of Faulkner's work explores race, it does so from a White perspective. There is little room for characters who are not white, and even less room for characters like Sutpen's first wife, Charles Bon, and Clytemnestra who have Black and White ancestry. They only constitute `r percent(black_white_all$percent[6])` of Yoknapatawpha's larger social world and `r percent(char_count_AA$percent[4])` in *Absalom, Absalom!*. All of which is to say that as far as demographics go, there are really only three types of characters Black, White, and to some extent Native Americans, everyone else is an exception to this rule. 


```{r weight_count}
weight_count <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  distinct(EventID, Race, event_words) %>%
  group_by(Race) %>%
  summarize(word_total = sum(event_words)) %>%
  mutate(percent = word_total / sum(word_total))
```

This picture changes quite radically, however, when character presence is taken into consideration. Using the event data, it is possible to calculate how often a character is present and for how long. This is quite a different way to look at character data. After all, if someone is a demographically unique character this may not necessarily indicate that the character is unimportant to the plot. For example, there is only one Extra Terrestrial character in the movie *E.T.*, but that character nonetheless gets a lot of "screen time." By that same token, we can calculate the "screen time" for each different race group in *Absalom, Absalom!*. Quite revealing is the fact that even though they are a demographic minority characters with heterogeneous ancestry are present in `r percent(weight_count$percent[3])` of the text and are thereby the second most present characters after White characters (`r percent(weight_count$percent[5])`). One way to read this predominance of these characters is Faulkner's attempt to trouble the color-line. By collapsing the ostensibly natural boundary between "Black" and "White", Faulkner exposes the very constructedness of those boundaries. Charles Bon is able to win the affections of Judith and Henry by passing as white, and only when his true origins are exposed does he suddenly become a threat. Likewise, his son with the unnamed "Octoroon" mistress, Charles Etienne Saint-Valery Bon, also crosses the color line with a woman pejoratively described as "blank". It would appear that Faulker is exposing how the color line operates. 

```{r}
total_present_length <- absalom_events_short %>% 
                        filter(PresentMentioned == "Present") %>% 
                        distinct(EventID, .keep_all = TRUE) %>% 
                        summarise(total_words = sum(event_words))

character_appearances <- absalom_events_short %>% 
                           filter(PresentMentioned == "Present") %>%
                          #distinct(CharacterName, .keep_all = TRUE) %>%
                          group_by(CharacterName) %>% 
                          summarize(word_total = sum(event_words)) %>% 
                          mutate(percent = word_total / total_present_length$total_words[1]) %>% 
                          filter(CharacterName %in% c("Clytemnestra", "Charles Bon", "Charles Etienne Saint-Valery Bon", "Thomas Sutpen")) %>% 
                          arrange(CharacterName)
                          

```

A less generous perspective might be that Faulkner is not so much concerned with destabilizing the color-line, but rather what racial segregation means for Whiteness. That is, if he were truly concerned with how the color-line affected the white and black communities, Charles Etienne's journey would be more pronounced. As it stands, he only appears in `r percent(character_appearances$percent[2])` of events, while Clytemnestra appears in slightly more (`r percent(character_appearances$percent[3])`). This is in contrast to Charles Bon who receives far more attention to both combined (`r percent(character_appearances$percent[1])`). The racial anxiety runs decidedly one way, men with possible Black blood passing as White. Clytie is not a threat and Charles Ettiene is more of a curiosity. 
What is more, all the characters are overshadowed by Thomas Sutpen (`r percent(character_appearances$percent[4])`). 

Sutpen's domineering presences redirects the racial question back to him. That is, why did his grand design fail. While the source of his folly can be traced back to a fundamental misunderstanding regarding racial cues: he does not recognize his first wife as part Black. There is a way in which the text makes an subtle and uncomfortable argument that part of the issue is that Sutpen does not embody the proper type of Whiteness. He is of low breeding and therefore can never hope to permanently establish a Plantation and record his name among the fine families of Yoknapatawpha: Sartoris, De Spain, Compson. His breeding is too low and his ambitions too grand for the county. WHiteness is as much a class issue as it is an understanding of race. 






With each narrator the social kaleidoscope changes and a new pattern is revealed. The new pattern likely says more about the narrator than it does about the actual "truth" of the matter, such as it is. 


### Character Count Table



```{r}
#fig <- plot_ly(char_count, labels = ~Race, values = ~n, type = 'pie')
#fig
```



```{r plot_weight_count}
fig <- plot_ly(weight_count, labels = ~Race, values = ~word_total, type = 'pie')
fig

```


### Character Pie Chart by Race


```{r mixed_race_characters}

weight_count_mx_race <- absalom_events_short %>% 
                filter(Race == "MixedBlackWhite") %>% 
               distinct(EventID, CharacterName, event_words) %>% 
                group_by(CharacterName) %>% 
                  summarize(word_total = sum(event_words)) %>% 
                  top_n(3)
```

```{r plot_mixed_race_characters}
fig <- plot_ly(weight_count_mx_race, x = ~CharacterName, y = ~word_total, type = 'bar')
fig

```
```{r all_characters}

weight_count_all_char <- absalom_events_short %>% 
                #filter(Race == "MixedBlackWhite") %>% 
               distinct(EventID, CharacterName, event_words) %>% 
                group_by(CharacterName) %>% 
                  summarize(word_total = sum(event_words)) %>% 
                  top_n(10)
```

```{r plot_mixed_race_characters2}
fig <- plot_ly(weight_count_all_char, x = ~CharacterName, y = ~word_total, type = 'bar')
fig

```


```{r narrator_break_up}

narrators <- absalom_events_short %>% 
              distinct(EventID, narrator, event_words) %>% 
              separate_rows(narrator, sep=",") %>% 
              group_by(EventID, event_words) %>% 
              arrange(EventID,narrator) %>%
              
              summarize(narrator = str_c(narrator, collapse = ",")) %>% 
              mutate(narrator = ifelse(str_count(narrator, ",")>0, "Multiple-Sources", narrator)) %>% 
                group_by(narrator) %>% 
                  summarize(word_total = sum(event_words)) %>% 
                  mutate(narrator = ifelse(word_total<1910, "Other", narrator)) %>% 
  mutate(narrator = str_trim(narrator))
              
```



```{r}
fig <- plot_ly(narrators, labels = ~narrator, values = ~word_total, type = 'pie')
fig

```

```{r}
narrators_single <- absalom_events_short %>% 
              distinct(EventID, narrator, event_words) %>% 
              separate_rows(narrator, sep=",") %>% 
              group_by(EventID, event_words) %>% 
              arrange(EventID,narrator) %>%
              add_count(EventID, name = "event_narrators") %>% 
              mutate(event_words = event_words/event_narrators) %>% 
              group_by(narrator) %>% 
               mutate(narrator = str_trim(narrator)) %>% 
              summarize(word_total = sum(event_words)) %>% 
                mutate(word_total = as.integer(word_total)) %>% 
              mutate(narrator = ifelse(word_total<3675, "Other", narrator))



```

```{r}
fig <- plot_ly(narrators_single, labels = ~narrator, values = ~word_total, type = 'pie')
fig
```

