---
title: '*Absalom, Absalom!* and the Digital Humanities: Charting Characters, Events, and Language'
author: "Johannes Burgers"
output:
  html_document:
    theme: yeti
    df_print: paged
  word_document: 
    reference_docx: citation_management/absalom_template.docx
bibliography: citation_management/absalom_bibliography.bib
csl: citation_management/modern-language-association.csl
always_allow_html: true
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, error=FALSE, tidy='styler')
knitr::knit_engines$set(python = reticulate::eng_python)
```

```{r load_packages}
#Data wrangling
library(tidyverse)
library(tidytext)
library(stringr)
library(lubridate)

#Python interoperability
library(reticulate)

#Visualization
library(plotly)

#Output
library(shiny)
library(scales)
library(webshot)
```

```{r load_text}
#Read in the text
absalom_txt <- as.data.frame(read_file("protected_data/absalom_cleaned_to_match_4_30_rev4.txt"))
colnames(absalom_txt) <- "text" 

absalom_df <- absalom_txt %>% 
mutate(text = gsub("[‘’]", "'", text)) %>% 
              unnest_tokens(text, text) %>% 
              summarize(text = str_c(text, collapse = " "))
```

```{r _load_event_sentences}
absalom_event_sentences <- read_csv("protected_data/absalom_event_sentences_clean_2021_5_10.csv")

#This file was created by recomposing the text based on the "First Words" column and matching the text segments with each event using a copy of the text. As this process is quite involved it has been branched out into a separate file: absalom_sentiment_text_processing. Full details on the process can be viewed there.

```

```{r global_variables}
absalom_length <- str_count(absalom_df$text, "\\S+")
```

```{r absalom_event_word_count}
event_length <- absalom_event_sentences %>%
  mutate(event_words = str_count(event_sentence,  "\\S+")) %>%
  select(EventID, begin_word, end_word, event_sentence, event_words)
```

```{r read_all_absalom_events}
absalom_all_events <- read_csv("data/absalom_all_events_2021_5_11.csv")
```

```{r event_dates}
aa_event_dates <- read_csv("data/absalom_event_dates_05_15.csv")

aa_event_dates <- aa_event_dates %>% 
                  mutate(StartDate=str_trim(StartDate)) %>% 
                  mutate(EndDate= str_trim(EndDate)) %>% 
                  select(EventID, StartDate, EndDate) %>% 
                  drop_na()
```

```{r chapter_numbers}
chapter_numbers <- read_csv("data/chapter_numbers.csv")
```

```{r clean_up_absalom_events}
#The full table is long and hard to work with.
absalom_events_short <- absalom_all_events %>%
  select(3:8, 11, 13, 21:28, 34, 50, 52:54) %>%
  mutate(narrator = str_extract(Summary, "(?<=(SOURCE)).+[^]</p>]")) %>%
  mutate(narrator = str_replace_all(narrator, ":", "")) %>%
  left_join(event_length, by = "EventID") %>%
  left_join(chapter_numbers, by = "PageNumber") %>%
  left_join(aa_event_dates, by = "EventID") %>%
  relocate(StartDate, .after = NarrativeStatus) %>%
  relocate(EndDate, .after = StartDate)
```

```{r read_all_DY_events}
dy_events <- read_csv("protected_data/full_database_2021_5_11.csv")
```

```{r clean_up_all_events}
dy_events_short <- dy_events %>%
  select(1:11, 21:27, 50, 52:54)
```

```{r load_chart_styling, echo=FALSE}
#Page styling

#This is where you create your own custom color palette for the traces.
faulkner_colorway = c("#132C53","#F27A18","#ae0700","#79b473","#38726c","#76bed0","#6b2d5c","#448b2d","#e6d812")

faulkner_colorway_bw = c("#999999", "#777777",	"#555555", "#333333",	
"#111111")

#This controls the background color for the entire chart. Probably best left white.
faulkner_paperbackground = c('rgba(255,255,255,0)')

#This controls the background for the plot. Probably best left white.
faulkner_plotcolor = c('rgba(255,255,255,.3)')

#Margin 

m <- list(l = 50, r = 50, b = 50, t = 50, pad = 4)

#Caption Style

fig_caption <- "font-family: 'Open Sans','Helvetica Neue',Helvetica,Arial,sans-serif; font-weight: normal; font-size:90%"

plot_font <- list(
  family = "'Playfair Display','Helvetica Neue',Helvetica,Arial,sans-serif",
  size = 12,
  color = '#363636')
```

## Introduction

*Absalom, Absalom!* is an unabashedly difficult text. Its structure is overwrought, its characters are inscrutable, and the prose is nearly impenetrable. There is also no shortage of digressions -- parenthetical, italicized, and otherwise -- that force even the most focused reader to re-read the same passage several times. Early on, in what could serve as a description of the narrative writ large, Mr.Compson refers to the dramatis personae of the Sutpen tragedy as a recovered and barely legible chemical formula that appears to be inert. When you bring all the characters together "nothing happens," and so you "re-read, tedious and intent, poring, making sure that you have forgotten nothing, made no miscalculation; you bring them together again and again nothing happens" (80). This sensation of mixing a formula that produces no reaction might feel familiar to some first-time readers. Perhaps the formula feels inert because the science is all wrong. The novel is, if anything, alchemical. Its power comes from its gnostic devotion to the *why,* which overshadows the more mundane matter of the *what*.

Like Sutpen's own divaricated progenerations, the text's difficulty has engendered multiple branches of critical interpretation. Although divergent in their aims and illuminating in their own ways, they are unified in their approach. Each relies on a methodology that builds evidence for a particular interpretation through a close-reading, supported by varying degrees of biographical, cultural, and historical context. As an alternative and complement to this approach, this essay deploys a methodology broadly described through the catch-all term: digital humanities. Generally, digital humanities scholars leverage the computational power of digital technologies to perform analyses that would be very difficult to do using traditional humanities methods. In this vein, the work here draws on the database and digital tools available through *Digital Yoknapatawpha* (*DY*). Hosted by the University of Virginia, *DY* is a resource for students and scholars of William Faulkner. The team has encoded all of the characters, locations, and events in Faulkner's Yoknapatawpha fictions. Through the site, users can study "deep atlases" of each text, explore archival material, listen to audio recordings of Faulkner, perform powerful relational searches, and analyze the texts using advanced data visualizations.

Using *DY* and its rich data sets, it is possible to investigate *Absalom, Absalom!* from multiple perspectives simultaneously. While it might be tempting to use the database to get to know the "facts" about the Sutpen family history, such an approach glosses over the highly interpretive nature of textual encoding, and, more to the point, contradicts *Absalom, Absalom!*'s ambivalence about ever truly knowing a story. Instead, a more productive approach is to use the database to showcase the multiplicity of interpretations the text permits. In this spirit, this essay has not been organized around a central thesis that makes a definitive claim about *Absalom, Absalom!*; instead it is a triptych of loosely connected vignettes structured around different data visualizations. The visualizations have been modified for the purposes of print; the full interactive versions are available online as a supplement. Each mini-essay covers a separate facet of the novel: characters, locations, events, and language.[^1] While each vignette could, arguably, be read separately, the themes echo each other: characters and race, events and history, language and storytelling. These themes are still as resonant today as they were in 1936. Indeed, current conversations about the racial legacy of America's past are as much about history's refusal to stay silent, as they are about finding the language to speak about that legacy in the present. *Absalom, Absalom!* suggests that those discussions do not end, but are, in fact, an end in and of themselves.

[^1]: All of the data was generated in the R programming language using the tidyverse suite of packages for the calculations and the plotly library for the graphics. The full repository is available at <https://github.com/joostburgers/absalom_four_dimensions> Due to copyright issues the repository does not include the *Absalom, Absalom!* text file used for data analysis. The text used was the 2011 Vintage Edition (Faulkner, *Absalom, Absalom!*), which corresponds with the 1990 edition we used for *DY*.

## Part 1: Characters

### "money, a house, a plantation, slaves, a family --- incidentally of course, a wife"

A well-worn quip about *Absalom, Absalom!* is that it is not a whodunit, but rather a why-done-it. Within the first several pages, the basic facts of the case are laid out: A stranger, Thomas Sutpen, appears in Jefferson one day, builds a plantation, and marries Ellen Coldfield. They have two children: Judith and Henry. Later, Henry's close friend, Charles Bon, wants to marry Judith, but on the day of their wedding, Henry shoots him and disappears. His daughter unwed, his son gone, and no heir to his estate, Sutpen slides into dissolution. Quentin Compson, the novel's main detective of sorts, will spend the remainder of the narrative trying to understand why Henry shot Bon by working through the various possible motivations in his conversations with the story's other narrators: Ellen's sister Rosa, his father, and his Harvard roommate, Shreve McCannon. No verifiable answer is ever provided, but Quentin avers that Charles was, in fact, Sutpen's son by a Haitian-born woman with part Black ancestry, Eulalia Bon. The resulting union between Judith and Charles would have constituted incest and "miscegenation" -- reproduction between a White and Black person. While there is a universal incest taboo, Quentin speculates that "it's the miscegenation" Henry "cant bear" (285). In probing the Sutpen family's fractious history, Quentin exposes the deep contradictions and outright hypocrisy that governed social relationships in the American South before and after emancipation. There was a strict system of formal and informal laws that maintained an inviolable color line between White and Black, but the lives and, indeed, genealogies of White and Black people could be deeply intertwined. In such a society, Charles Bon could be both brother and Other to Judith.

*Digital Yoknapatawpha* provides powerful tools for investigating these social relationships. Users can pull up the biographies of individual characters, and explore interactive genealogies of each major family, including the Sutpen's [@massariYoknapatawphaFirstFamilies]. Taking the data, it is also possible to aggregate characters by group and create a demography of the world of *Absalom, Absalom!*. This demography reveals how racial divisions in the South did not exist outside of the social norms surrounding gender, sexuality, and class, but, instead, were constituted through them. Thus, while the text troubles social distinctions based on race, it often does so by re-inscribing certain assumptions about gender, sexuality, and race.

```{r black_white_ratios}
all_characters <- dy_events_short %>%
distinct(CharacterName) %>%
nrow()

black_white_all <- dy_events_short %>%
# filter(SourceTextCode!="AA") %\>%
distinct(CharacterName, .keep_all = TRUE) %>%
count(Race) %>%
mutate(percent = n / sum(n))
```

```{r count_char}
char_count_AA <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>% 
  distinct(CharacterName, Race) %>%
  count(Race)  %>%
  mutate(percent = round(n / sum(n),2)) %>% 
 mutate(Race = ifelse(Race=="MixedBlackWhite", "Mixed Black White", Race))
```

The demographics of Faulkner's fictional Yoknapatawpha differ in slight but important ways from that of the actual Lafayette County. Don Doyle's thorough historical account of northern Mississippi states that the ratio of Black (40%) and White (60%) people in the county was relatively stable throughout Lafayette's history [-@RN16 303]. On the map he drew for *Absalom, Absalom!*, Faulkner flips this ratio and claims that there are 6,298 (40%) Whites and 9,313 (60%) "Negroes" [@railtonFaulknerMappingMapping]. These proportions are not reflected in his fictions, which are overwhelmingly White. There are `r comma(all_characters)` characters, `r comma(black_white_all$n[10])` (`r percent(black_white_all$percent[10])`) of whom are White. There are only `r comma(black_white_all$n[2])` (`r percent(black_white_all$percent[2])`) Black and `r comma(black_white_all$n[4])` (`r percent(black_white_all$percent[4])`) Native American characters. That said, Faulkner does carve out space for people who complicate these strict racial categories. Characters of mixed Black and White ancestry constitute `r percent(black_white_all$percent[6])` of Yoknapatawpha's larger social world and `r percent(char_count_AA$percent[3])` of the population in *Absalom, Absalom!*. Although certainly small, this number is remarkably consistent with the available census data, even if, as Doyle points out, the number of people recorded as "mulatto" was surely higher because of the "secrecy and denial that surrounded liaisons across racial lines" [-@RN16 148].

Nevertheless, this static picture of the characters is deceptive for several reasons. First, it masks the editorial decisions behind the data. For example, there is no concrete evidence that Charles Bon's mother was part Black, and it is therefore not self-evident that he should be coded as "Mixed Black White." At the same time, since so much of the novel revolves around this possibility, the editors chose this as the best option. Second, the raw demographic counts also do not consider how frequently certain characters appear in the text. This is calculated by counting the number of times a specific character or types of characters appear in events across the text. [^2] This measure is comparable to how much "stage time" a character might get in a play. The distinction is significant. For instance, Othello is the only Moor character in *Othello*, but he is very much present on stage. The chart below (Figure 1) shows both the raw demographic data and the frequency of character presence by race.

[^2]: The data for this essay is slightly more nuanced than for other texts in *DY*, because in conjunction with event frequency the event length by word count was also taken into consideration. That being said, there are multiple ways to calculate character "presence" that result in divergent absolute differences between numbers, but their relative share -- more, less, least -- stays fairly consistent.

```{r weight_count}
weight_count <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
 # distinct(EventID, Race, .keep_all = TRUE) %>% 
  group_by(Race) %>%
  summarize(word_total = sum(event_words)) %>%
  mutate(percent = round(word_total / sum(word_total),2)) %>% 
  mutate(Race = ifelse(Race=="MixedBlackWhite", "Mixed Black White", Race)) %>% 
  filter(percent >0)
  



```

```{r plot_demography, fig.cap="Figure 1: Demography by raw count and by character presence. The image on the left shows what percentage of each type of character is present in the text. The image on the right shows how often they appear. Note that \"Indian\" and \"Multiracial Groups\" dissappear, because they are present in less than 1% of the text."}


plot_demography <- plot_ly()

plot_demography <- plot_demography %>% 
      add_pie(
      values =  ~char_count_AA$n,
    labels = ~ char_count_AA$Race,
    type = 'pie',
    name = "Character Count",
    text = paste(char_count_AA$Race, char_count_AA$percent*100, "%"),
    textinfo = 'text',
    domain = list(x=c(.1,.45), y = c(.1,.9))
   
  )

plot_demography <- plot_demography %>% 
  add_pie(
        values = ~ weight_count$word_total,
    labels = ~ weight_count$Race,
    type = 'pie',
    name = "Character Appearance",
    text = paste(weight_count$Race, weight_count$percent*100, "%"),
    textinfo = 'text',
    domain = list(x = c(.55,.9), y = c(.1,.9))
 
  )

plot_demography <- plot_demography %>% layout(
    title = list(text = "Character Demography"),
    font = plot_font,
    showlegend = FALSE,
    colorway = faulkner_colorway_bw,
    margin = m,
    autosize = F,
    width = 1024
    
  )

plot_demography <- plot_demography %>% add_annotations(
  x = c(.2,.8),
  y = c(1, 1),
  xref = 'paper',
  yref = 'paper',
  showarrow = FALSE,
  text = c("Raw Count", "Present in Events"),
  font = list(size = 14),
  align = 'left'
  
)

plot_demography <-  plot_demography %>%
  config(
    toImageButtonOptions = list(
      format = "png",
      width = 1000,
      heigth = 4000
    )
  )

  
plot_demography
```

The charts highlight the prominent role of mixed ancestry characters. Even though they are a demographic minority, the relative share of mixed ancestry characters increases to `r percent(weight_count$percent[2])` when their textual presence is taken into account. The statistics affirm the obvious: characters with mixed ancestry feature prominently. What is harder to detect without a statistical overview is that the attention to mixed ancestry characters does not reduce the focus on White characters, whose relative representation actually increases. Black characters are the ones forced into the background. This nuances how Faulkner explores miscegenation in this text. Speaking of the culture at large, Eric Sundquist explains that "The children of miscegenation not only threatened to blur the distinctions that made slavery possible, but the unions that produced them threatened to distort sexual relations and marriage itself, for black and white alike, into peculiar and tragic forms" [-@sundquistFaulknerHouseDivided1983 116]. The demography suggest that Faulkner is very much concerned with how interracial liaisons trouble Whiteness, but does not explore what they might mean for Blackness.

```{r mixed_ancestry_present}
total_present_length <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  distinct(EventID, .keep_all = TRUE) %>%
  summarise(total_words = sum(event_words))

character_appearances <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
    group_by(CharacterName, Race) %>%
  summarize(word_total = sum(event_words)) %>%
  mutate(percent = word_total / total_present_length$total_words[1]) %>% 
  filter(
    CharacterName %in% c(
      "Clytemnestra",
      "Charles Bon",
      "Charles Etienne Saint-Valery Bon",
      "Jim Bond",
      "Mrs. Charles Bon|Octoroon Mistress",
      "Mrs. Thomas Sutpen(1)"
    )
  ) %>%
  arrange(CharacterName)
```

This is not entirely surprising. With the exception of Charles, all of the main characters are White, and Faulkner uses the central dynamic between Sutpen, Bon, Henry, and Judith to complicate the color line, something other scholars have covered quite extensively.[^3] Still, only looking at the main characters in a text can be distorting. One of the democratizing features of *DY* is that it renders minor characters more visible. This perspective reveals that the representation of race is heavily mediated by gender and sexuality. Charles Bon, as a man with ostensibly Black blood who wants to marry a White woman, receives as much attention (`r percent(character_appearances$percent[1])`) as all the other mixed ancestry characters combined -- Eulalia Bon (`r percent(character_appearances$percent[6])`), "octoroon" mistress (`r percent(character_appearances$percent[5])`), Clytemnestra (`r percent(character_appearances$percent[3])`), Charles Etienne (`r percent(character_appearances$percent[2])`), and Jim Bond (`r percent(character_appearances$percent[4])`). This disparity is curious because the text is full of racial transgressions. Eulalia Bon, Sutpen's first wife, appears infrequently, even though she commits a similar "blood" crime, and helps precipitate the failure of his design. Likewise, Charles's "octoroon mistress" only exists to fulfill his desires. Clytemnestra -- the daughter of an enslaved woman who was sexually exploited by Sutpen -- suffers a singularly cruel fate. While Davis offers an insightful and redemptive reading of Clytie as the "felt black presence that pervades the South" [-@davisSignifyingAbstractionReading2003 81], Sutpen's mixed ancestry daughter does not, statistically at least, match the narrative presence of her half-brother, Bon. She is also continuously disassociated from the family, and is alternately described as "negro half sister" (87), "wild" (126), "without sex or age" (109), "negress" (139), or, simply, "servant" (149). Of the mixed ancestry women, she is able to be more present because she has been stripped of all femininity, sexuality, and, indeed, agency. Her one direct intervention in the family saga is also her final one: burning down the house. By virtue of their race, *Absalom, Absalom!'s* mixed ancestry women cannot be loving wives or caring mothers, only useful servants.

[^3]: One of the first and most extensive studies is Eric @sundquistFaulknerHouseDivided1983's *Faulkner: The House Divided*. His early intervention pointed to Faulkner's treatment of miscegenation as central to understanding his work. Barbara @laddDirectionHowlingNationalism1994's "The Direction of the Howling": Nationalism and the Color Line in *Absalom, Absalom!*" explores how the figure of the creole who passes as white, Charles Bon, represents a fundamental threat to white southerners conception of their own history and possible futures. Another seminal piece is that of Thadious @davisSignifyingAbstractionReading2003 who demonstrates that the narrators each project a different abstraction of "The Negro" in order to reconcile their own relationship with the past. Other more recent relevant scholarship includes that of @crawfordInterracialEmbodimentContradiction2004; @sugimoriRacialMixtureRacial2008; @cardenFatherlessChildrenPostPatrilineal2013; and @sneadFiguresDivisionWilliam2017.

Meanwhile, men of mixed ancestry are also portrayed as embodying differences in gender and sex. Bon's son by his "octoroon" mistress, Charles Etienne is highly feminized with "light bones and womanish hands" (162). His marriage to a woman disparagingly described as "coal black and ape-like" (166) leads to violent altercations with both Black and White men who believe him to be White, with some even claiming that he suffers from a "sexual perversion" (167). Conversely, their son, Jim Bond, is a "hulking slack-mouthed" brute (173). He is so hyper-masculine that he is incapable of rational thought, and cannot express any desire at all except unintelligible howling. These various alternative versions of mixed ancestry all suggest that in the minds of the narrators, and perhaps Faulkner as well, racial divergence corresponds to gender and sexual Otherness. It also implies that gender and sexual difference were markers of racial impurity. In this sense, the real danger is not that Charles desires Judith, but rather that she desires him.

```{r sutpen_social}

Sutpen_events <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  filter(CharacterName == "Thomas Sutpen")

Sutpen_unique_all <- absalom_events_short %>%
  semi_join(Sutpen_events, by = "EventID") %>%
  filter(PresentMentioned == "Present") %>%
  filter(!CharacterName %in% c("Thomas Sutpen","Judith Sutpen", "Henry Sutpen", "Charles Bon", "Clytemnestra")) %>% 
  group_by(Class) %>% 
  summarise(total_words = sum(event_words)) %>% 
  mutate(percent = total_words/sum(total_words)) %>% 
  arrange(Class)

Sutpen_with_upper_class <- absalom_events_short %>%
  semi_join(Sutpen_events, by = "EventID") %>%
  filter(PresentMentioned == "Present") %>%
  filter(!CharacterName %in% c("Thomas Sutpen","Judith Sutpen", "Henry Sutpen", "Charles Bon", "Clytemnestra")) %>% 
  filter(Class =="Upper Class") %>% 
  group_by(LocationTitle) %>% 
  distinct(EventID,.keep_all = TRUE) %>% 
  summarise(total_words = sum(event_words)) %>% 
  mutate(percent = total_words/sum(total_words))

```

Any discussion of race would be incomplete without also considering class. In part, Sutpen's "compelling dream" fails because of he is not the right type of White (135). Born into humble circumstances in the mountains of western Virginia, Sutpen is unfamiliar with social distinctions. Only when he moves closer to the plantations of Virginia does he realize that there were differences not only between "white men and black ones," but also "between white men and white men" (183). When he is barred from entering the front door of the Pettibone plantation by an enslaved Black servant, it reveals the divide between him and upper-class White men. His initial exclusion from the elite, becomes the source of his downfall. Had he been born into status, there would be no need to travel to Haiti to become wealthy, and he would have married into a bloodline that would have been guaranteed to be "pure" White.

Sutpen never achieves the class status, and, by extension, Whiteness to which he aspires. Despite being the richest planter in the county, Mr. Compson points out that Sutpen's "flowering was a forced blooming" and that he was "playing the scene to an audience" (57). Unable to relinquish his roots, he surrounds himself with characters outside of high society. If his immediate family is excluded, Sutpen spends most of his time with: enslaved Blacks (`r percent(Sutpen_unique_all$percent[1])`), the captive French architect and his middle-class in-laws (`r percent(Sutpen_unique_all$percent[7])`), and lower-class and "poor whites" (`r percent(Sutpen_unique_all$percent[9]+Sutpen_unique_all$percent[6])`). Even when he is among the upper-class (`r percent(Sutpen_unique_all$percent[10])`), it is always in a grotesque pantomime of class status: brawling with the people he enslaved (21), his "grand" wedding to Ellen (43), or the "hunting party" for the French architect.

Sutpen's "poor white" background forecloses the possibility of him establishing himself permanently among the ranks of the elite families of Yoknapatawpha - Sartoris, De Spain, Compson. Class, like race, is inherited. The paradox, of course, is that in Northern Mississippi many of the planters who arrived in the 1830s, as Sutpen does, were not "born" into their position; they were hardscrabble men of few scruples who masked their violent rise from indigence through a veneer of gentility [@raileyNaturalAristocracyHistory1999b 118-19]. Historically speaking, the differences between Sutpen and the plantocracy of the town, is a difference in degree, but not in kind, even if in the world of Yoknapatawpha certain families like Sartoris and Compson are simply better than others, and, in a sense, Whiter. By that same token, there are certain expressions of gender and sexuality that are Whiter than others, and therefore sanctioned. The text complicates the inviolability of some of these distinctions, but the dynamics of the demographic data suggest that the narrators, and perhaps the author as well, are not always able to look beyond the entrenched social categories of their own time.

```{r death_AA}
deaths_AA <- absalom_events_short %>% 
              filter(PresentMentioned=="Present") %>% 
              distinct(CharacterName, .keep_all = TRUE) %>% 
              count(CauseOfDeath) %>% 
              drop_na() %>% 
              mutate(percent = n/sum(n)) %>% 
              arrange(CauseOfDeath)

total_deaths <- sum(deaths_AA$n)

old_age <- absalom_events_short %>% 
              filter(PresentMentioned=="Present") %>% 
              distinct(CharacterName, .keep_all = TRUE) %>% 
              filter(CauseOfDeath =="Old Age")

```

```{r narrative_structure}
#This primes the data for plotly. TODO upstream this ina more logical way. This is just a grab bag of cleaning.

absalom_chronology <- absalom_events_short %>%
  distinct(EventID, .keep_all = TRUE) %>%
  select(1:8, 10, 28:29) %>%
  mutate(Summary = str_remove(Summary, '<p>')) %>%
  mutate(Summary = str_remove(Summary, '</p>')) %>%
  mutate(Summary = gsub("((\\S+\\s){7})", "\\1<br> ", Summary)) %>%
  arrange(`Chronological Order`) %>%
  mutate(Chrono = row_number())  %>%
  mutate(StartDate = as_date(StartDate)) %>%
  arrange(OrderWithinPage) %>%
  mutate(OrderPageEnd = lead(
    OrderWithinPage,
    n = 1,
    default = last(OrderWithinPage)
  )) %>%
  mutate(EventLength = OrderPageEnd - OrderWithinPage) %>%
  relocate(OrderPageEnd:EventLength, .after = OrderWithinPage) %>%
  mutate(NarrativeStatus = str_replace_all(NarrativeStatus, "\\+", ""))

absalom_chronology_bw <-  absalom_chronology %>%
  pivot_wider(names_from = "NarrativeStatus",
              values_from = Chrono)


absalom_jump <- absalom_chronology %>%
  mutate(Chrono2 = lag(Chrono, default = Chrono[1])) %>%
  mutate(difference = Chrono - Chrono2) %>%
  mutate(
    jump = case_when(
      difference < 0 ~ "flashback",
      difference > 1 ~ "flashforward",
      difference <= 1 ~ "linear"
    )
  ) %>%
  count(jump) %>%
  mutate(percent = n / sum(n))

narrative_present <- absalom_chronology %>%
  mutate(StartDate = as_date(StartDate)) %>%
  mutate(temporality = ifelse(StartDate >= StartDate[1], "Narrative Present", "Past")) %>%
  group_by(temporality) %>%
  summarise(total_words = sum(event_words)) %>%
  mutate(percent = total_words / sum(total_words))

absalom_status <- absalom_chronology %>%
  group_by(NarrativeStatus) %>%
  summarise(total_words = sum(event_words)) %>%
  mutate(percent = total_words / sum(total_words))
```

## Part 2: Events

### "Maybe happen is never once"

One of the most challenging parts of *Absalom, Absalom* is certainly its narrative structure. Readers familiar with Faulkner's other works might recognize the experimentation with different perspectives, stream-of-consciousness narration, and the continuous revision of basic plot points that were also features of *The Sound and the Fury*, *As I Lay Dying*, and *Light in August*. In some respects, *Absalom, Absalom!* represents a synthesis of these narrative innovations. *Absalom, Absalom!* features multiple narrators whose stories all flow into Quentin's stream-of-consciousness narration, which makes them, at times, indistinguishable. What is more, the different narrators continuously revise major plot points, and to borrow Mr. Compson's apt description, drop these "into the telling as you might flick the joker out of a pack of fresh cards without being able to remember later whether you had removed the joker or not" (203). Disentangling the text is therefore equal parts exasperating and exhilarating.

To complicate matters further, the narrative is decidedly non-linear. Only `r percent(absalom_jump$percent[3])` of the events follow a sequential order, the remainder of the plot either jumps forward (`r percent(absalom_jump$percent[2])`) or backwards (`r percent(absalom_jump$percent[1])`) non-sequentially. These jumps might appear random, but are actually part of a deep structure. Faulkner's notes for *Absalom, Absalom!* make it clear that he had a core design he kept elaborating [@blotnerFaulknerBiography2005d 347; @RN33 11]. Because the text is so confusing, deciphering its structure has been somewhat of a cottage industry [@RN31; @RN29 201-02; @RN32], but it was not until the assiduous work of Stephen Railton in 2003 that it became possible to visualize its chronology on an interactive timeline.[^4] This data formed the foundation for encoding *Absalom, Absalom!'s* chronology in the *DY* database, which, in turn, made it is possible to generate a plot structure chart.[^5]

[^4]: Unfortunately, the visualization is no longer operational, but efforts are being made to revive it.

[^5]: For an extensive and rather technical overview of this process see: @burgersImaginingContinuouslyPresent2021

Conceptually, a text can have two different orders: story and plot. The story is the sequence of events as they happened (chronology), and the plot is the order in which they are told (page order).[^6] A plot structure chart places story/chronology on the y-axis and plot/page order on the x-axis.[^7] Events that happen later chronologically are higher on the y-axis, while events that happen earlier are lower. Meanwhile, event that are told in the beginning are on the far-left of the x-axis, and those told at the end are on the far-right. If the story is told in the sequence of events as it occurs, it will rise at a 45$^\circ$ slope. Most of *Absalom, Absalom!* is out of order and deviates from this 45$^\circ$ angle. For example, Sutpen's origins in western Virginia are not revealed until chapter 7. Chronologically this is an early event, but in the plot it is told much later. This point is low on the y-axis, but about three-fourths of the way across on the x-axis. Generally, when events drop down the y-axis this is a sign of a flashback. Conversely, a jump up is a flashforward. In *DY*, the narrative structure of every Faulkner text can be accessed through the "Narrative Analysis" dashboard [@burgersNarrativeStructureAnalysis].

[^6]: This idea was originally developed by the Russian formalists in the 1920s, who labeled story and plot, fabula and syuzhet, respectively [@shklovskiiTheoryProse1990]. The distinction between the two elements was also independently intuited by E.M. Forster around the same time [@forsterAspectsNovel1927].

[^7]: Two sets of data are available for determining the event sequence in *DY*: date range and the chronological order of the event. The data range is a best estimate of when an event happened on a historical timeline. The chronological order takes all events in the novel and numbers them from 1 to the total number of events (i.e. 1, 2, 3...644). It is the most exact representation of the event order. Since the chart steps through all the events in sequence, the distance between all the events is 1, and any relative distance between them is removed. This level of abstraction helps provide insight into the plot structure irrespective of dates. Using the date range is less exact, but the advantage is that it shows the relative distances between events. Both sets of charts are available on the Narrative Structure Analysis dashboard.

The chart below shows the narrative structure of the text, and how that information is conveyed to the reader: *narrated* through the external narrator, *told* by one of the characters, or conveyed through one of the other types of narration identified by *DY*: *remembered*, *hypothesized*, or *narrated+consciousness*. The latter three were consolidated into one data series because their number was relatively low. The chart has also been annotated with major plot points.

```{r import_annotations}
#creates a clean csv for export. Not necessary

#chronology_annotations_export <- absalom_chronology %>%
#  select(4, 8:9, 11, 14)

#write_csv(chronology_annotations_export, "annotations_export.csv")

chrono_ann_import <- read_csv("data/annotations_import.csv")
chrono_ann_import_secondary <-
  read_csv("data/annotations_import_secondary.csv")

chrono_ann_import <- chrono_ann_import %>%
  drop_na(event)

```

```{r create_chapter_markers}
chapters_single <- absalom_chronology %>%
  distinct(chapter, .keep_all = TRUE) %>%                             select(OrderWithinPage, chapter)
```

```{r plot_aa_bw, fig.cap= "Figure 3: Plot Structure of <i>Absalom, Absalom!</i>. Particularly notable, is the repetition of an S pattern throughout. This indicates a similar structure across the chapters."}
plot_aa_bw <- plot_ly(
  absalom_chronology_bw,
  x = ~ OrderWithinPage,
  y = ~ Narrated,
  name = 'Narrated',
  type = 'scatter',
  mode = 'markers',
  marker = list(
    symbol = 'square',
    line = list(
      width = 1,
      color = "#000000",
      opacity = 1
    ),
    color = "#FFFFFF",
    opacity = .6
  ),
  hovertemplate = paste(
    'Page: %{x}',
    '<br>',
    'Narrative Status: Narrated',
    '<br>',
    'Chronological Order: %{y}',
    '<br>',
    'Start Date:',
    absalom_chronology_bw$StartDate,
    '<br>',
    'Summary:',
    '<br>',
    absalom_chronology_bw$Summary,
    '<extra></extra>'
  )
)

plot_aa_bw <- plot_aa_bw %>% add_trace(
  x = ~ OrderWithinPage,
  y = ~ Told,
  name = 'Told',
  type = 'scatter',
  mode = 'markers',
  marker = list(
    symbol = "circle",
    line = list(
      width = 1,
      color = "#000000",
      opactiy = 1
    ),
    color = "#D4D4D4",
    opacity = .6
  ),
  hovertemplate = paste(
    'Page: %{x}',
    '<br>',
    'Narrative Status: Told',
    '<br>',
    'Chronological Order: %{y}',
    '<br>',
    'Start Date:',
    absalom_chronology_bw$StartDate,
    '<br>',
    'Summary:',
    '<br>',
    absalom_chronology_bw$Summary,
    '<extra></extra>'
  )
)

plot_aa_bw <- plot_aa_bw %>% add_trace(
  x = ~ OrderWithinPage,
  y = ~ Remembered,
  legendgroup = "Other",
  name = paste('Remembered<br>Hypothesized<br>Narrated+Consciousness'),
  type = 'scatter',
  mode = 'markers',
  marker = list(
    symbol = "diamond",
    line = list(
      width = 1,
      color = "#000000",
      opacity = 1
    ),
    color = "#494848",
    opacity = .6
  ),
  hovertemplate = paste(
    'Page: %{x}',
    '<br>',
    'Narrative Status: Remembered',
    '<br>',
    'Chronological Order: %{y}',
    '<br>',
    'Start Date:',
    absalom_chronology_bw$StartDate,
    '<br>',
    'Summary:',
    '<br>',
    absalom_chronology_bw$Summary,
    '<extra></extra>'
  )
  
)
plot_aa_bw <- plot_aa_bw %>% add_trace(
  x = ~ OrderWithinPage,
  y = ~ Hypothesized,
  legendgroup = "Other",
  name = 'Other',
  showlegend = FALSE,
  type = 'scatter',
  mode = 'markers',
  marker = list(
    symbol = "diamond",
    line = list(width = 1,
                color = "#000000"),
    color = "#494848"
  ),
  hovertemplate = paste(
    'Page: %{x}',
    '<br>',
    'Narrative Status: Hypothesized',
    '<br>',
    'Chronological Order: %{y}',
    '<br>',
    'Start Date:',
    absalom_chronology_bw$StartDate,
    '<br>',
    'Summary:',
    '<br>',
    absalom_chronology_bw$Summary,
    '<extra></extra>'
  )
)
plot_aa_bw <- plot_aa_bw %>% add_trace(
  x = ~ OrderWithinPage,
  y = ~ NarratedConsciousness,
  legendgroup = "Other",
  name = 'Other',
  type = 'scatter',
  showlegend = FALSE,
  mode = 'markers',
  marker = list(
    symbol = "diamond",
    line = list(width = 1, color = "#000000"),
    color = "#494848"
  ),
  hovertemplate = paste(
    'Page: %{x}',
    '<br>',
    'Narrative Status: Narrated+Consciousness',
    '<br>',
    'Chronological Order: %{y}',
    '<br>',
    'Start Date:',
    absalom_chronology_bw$StartDate,
    '<br>',
    'Summary:',
    '<br>',
    absalom_chronology_bw$Summary,
    '<extra></extra>'
  )
)

plot_aa_bw <-  plot_aa_bw %>%
  layout(
    title = list(text = "Plot Structure Chart: <i>Absalom, Absalom!</i>"),
    xaxis = list(
      tickmode = 'array',
      tickvals = chapters_single$OrderWithinPage,
      ticktext = chapters_single$chapter,
      gridwidth = 1.5,
      title = list(text = "Chapter")
    ),
    font = plot_font,
    margin = m,
    yaxis = list(title = list(text = "Chronology"),
                 showgrid = FALSE),
    autosize = FALSE,
    width = 1024,
    height = 400
  )

plot_aa_bw <- plot_aa_bw %>%
  add_annotations(
    x = chrono_ann_import$OrderWithinPage,
    y = chrono_ann_import$Chrono,
    text = chrono_ann_import$event,
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 3,
    arrowwidth = 1,
    arrowsize = .5,
    ax = chrono_ann_import$xoffset,
    ay = chrono_ann_import$yoffset,
    font = list(color = chrono_ann_import$fontcolor),
    bgcolor = chrono_ann_import$bgcolor,
    borderpad = 0,
    borderwidth = 0
  )

plot_aa_bw <- plot_aa_bw %>%
  add_annotations(
    x = chrono_ann_import_secondary$OrderWithinPage,
    y = chrono_ann_import_secondary$Chrono,
    text = chrono_ann_import_secondary$event,
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 3,
    arrowwidth = 1,
    arrowsize = .5,
    ax = chrono_ann_import_secondary$xoffset,
    ay = chrono_ann_import_secondary$yoffset,
    font = list(color = "rgba(255,255,255,0)"),
    bgcolor = "rgba(255,255,255,0)",
    borderpad = 0,
    borderwidth = 0
  )

plot_aa_bw <-  plot_aa_bw %>% 
  config(
    toImageButtonOptions = list(
      format = "svg",
      width = 1024,
      height = 400
    )
  )

plot_aa_bw

```

At first blush, the above chart can be a bit overwhelming, and it is easiest to break it down into several components. First, the top line of squares that gradually rises from left to right represents the narrative present. These are all the conversations Quentin has with Rosa, his father, and Shreve during the summer of 1909 and winter of 1910. The major stretches of the story told by one narrator are also marked on this top line: Rosa, Mr. Compson, Quentin, Shreve, and, finally, Shreve and Quentin together. Below this top line are all the events that happened in the past. The gray dots appear to "bubble up" randomly, but they actually cluster into elongated "S-curves." The curve gives an indication of the narrative pacing. Each narrator starts and ends their story with a high density of events, but skips through the middle with larger chronological jumps. Some curves start farther down, and others end higher up. This difference in clustering makes more sense relative to the chapter markers. Moving from left to right, there is a small S at the bottom of chapter 1, a second one that stretches from chapter 2 to 3, the most substantial one that goes all the way from chapter 4 to the end of chapter 6, and then finally two individual curves in chapter 7 and 8. These patterns show how chapters 1-3 act as a preview, while chapters 4-6 develop the rising action. Meanwhile, chapters 7 and 8 each in their own way represent a climax, as information that was hinted at previously is explained more fully, if still incompletely. Within chapters there is a consistent pattern as well. They open in the present, drop down into the "deep" past, and work their way back up towards the present. There might be some errant events that float between the S-curve and the narrative present, but generally these are previews for events that will happen later. The pattern is pretty regular overall, and all the chapters work together to tell the "entire" story of Sutpen. With a little imagination, the curves slide into each other to make one long curve of the events in the right order. To help understand what is going on at each point on the chart, certain major events have been annotated. On occasion, some, though certainly not all, repetitions of major events are marked. This is particularly true of Henry shooting Bon, which is alluded to at least four times.

Based on the chart, it also possible to make basic inferences about patterns in the plot, in particular Faulkner's use of delayed revelation. In the beginning, Sutpen mysteriously appears in chapters 1 and 2, but we do not get his full story until chapter 7, quite some time after it is revealed that he dies at the end of Chapter 5. His cause of death only becomes clear by the end of Chapter 7. The story of Henry shooting Charles is told in the first couple of pages, and again in chapters 4 and 5, but we do not get the fuller version until the end of Chapter 8. The narrative gives glimpses of the past, but only provides their full context much later. To add to the complexity of the history, it is told in multiple different ways through different narrators. Generally, events in the past are "told," but they can also be narrated by the third person narrator, or use more esoteric forms of narration like hypothesizing or a memory within a telling.

Visually, it is useful to imagine the S-curves as continuous undulations of the past disrupting the placid suspension of the narrative present. The various narrators are not so much plumbing the past, but rather riding its currents to navigate the present, or as Quentin muses:

> Maybe nothing ever happens once and is finished. Maybe happen is never once but like ripples maybe on water after the pebble sinks, the ripples moving on, spreading, the pool attached by a narrow umbilical water-cord to the next pool which the first pool feeds, has fed, did feed, let this second pool contain a different temperature of water, a different molecularity of having seen, felt, remembered, reflect in a different tone the infinite unchanging sky, it doesn't matter: that pebble's watery echo whose fall it did not even see moves across its surface too at the original ripple-space, to the old ineradicable rhythm.... (210)

Following this metaphor, Sutpen's arrival is the pebble that ripples through the history of Yoknapatawpha. The narrators each describe the subsequent history in their own way, thereby causing the ripples to undulate and resonate at different frequencies, sometimes negating and sometimes amplifying one another, always disrupting the present. The patterns are all different, but distinguishing where one begins and another ends is impossible.

## Part 3: Language

### "it struck, word by word, the resonant strings of remembering"

Characters and events do not exist independently in a text, they are all embedded in language. To state the obvious, authors tell stories with words. Though *Absalom, Absalom!* is certainly a story, it is as much about the telling as it is about the tale. If Faulkner's prose can seem impenetrable, it is because it is. Yet, the impenetrability helps the language call attention to itself. To be sure, scholars have pointed out this effect throughout the long genealogy of Faulkner criticism, even if the theoretical idioms in which they do so offer different nuances [@brooksIncredulousNarrationAbsalom1982d 263; @hodgsonLogicalSequenceContinuity1971 103-104; @zoellnerFaulknerProseStyle1959 500]. Computational analysis amplifies these interpretations by visualizing how Faulkner's language resists the possibility of telling a story fully.

```{r sentence_length}

absalom_clean <- absalom_txt %>% 
mutate(text = gsub("[,\\!\\?\\.](?=[^’‘]*’[\\s—-])"," ",text, perl=TRUE))


#Regex explained:
#[,\\!\\?\\.] find any of !,?.
#(?=[^’‘] not followed by curly quote. The closing curly quote is counterintuitive here, but it helps exclude 's as in Sutpen's. Instead, we want all curly quotes followed with a white space \s: 
#*’[\\s—-]) the en and em dash are there because sometimes these can abutt the closing curly quote.

absalom_sentences <- absalom_clean %>%
  unnest_regex(word, text, "[.?!]") %>%
  mutate (string_length = str_count(word, "\\S+"))  

absalom_sentences_mean <-  absalom_sentences %>%
  summarize(mean_length = mean(absalom_sentences$string_length))

absalom_sentences_long <- absalom_sentences %>%
  filter(string_length > 99) %>%
  nrow()
```

The most infamous stylistic feature of *Absalom, Absalom!* is undoubtedly the length of the sentences. In fact, according to the 1983 *Guinness Book of World Records* it holds the dubious record of the longest sentence in the English language -- 1,300 words [@RN34 228]. The claim is absurd. James Joyce's *Ulysses* surely has longer sentences, and even Faulkner beat his own "record" quite handily with a 1,600 word sentence in *Go Down, Moses* [@RN35 102]. Faulkner's long sentences are more than a matter of trivia, however; they quite literally stretch the very concept of the sentence, and by extension the English language. Consequently, they are difficult to process for humans and computers alike. Normally, computing sentence length is a simple procedure. With *Absalom, Absalom!*, detecting the end of a sentence is complicated by the many parenthetical phrases and embedded quotations. The long sentences represent more than technical challenge though, they raise fundamental questions about the structure of language. There are no hard and fast rules as to how long a sentence can be, and yet eventually the structure becomes so complex that it might be syntactically correct but borders on semantic incoherence. Perhaps it is better to say, therefore, that the novel is a collection of clauses occasionally interrupted by punctuation. Given this caveat, the computed average sentence length of `r round(absalom_sentences_mean$mean_length,0)` words per sentence is at best a rough measure. What is clear though, is that this is significantly longer than other works in English written in the 1930s, which use only about 13 words per sentence [@RN37 233]. The average masks the sheer number of long sentences: `r absalom_sentences_long` sentences exceed 100 words.

For Faulkner, these long sentences are more than rhetorical pyrotechnics, they try to capture how the past continuously intrudes upon the present. Asked in an interview about why he wrote such long sentences, Faulkner contends that there is no such thing as "was" because everyone brings their past with them. His long sentences are a way to capture this sensation, "a man, a character in a story at any moment of action is not just himself as he is then, he is all that made him, and the long sentence is an attempt to get his past and possibly his future into the instant in which he does something" [@faulknerFaulknerUniversityClass1965 84]. In a sense, the prose complements the narrative structure, which also sees the past continuously returning.

```{r parentheticals}

#NOTE: The amount of parentheses in the text is uneven. An extra one was added at the end of chapter 6.
parentheses_file <-
      as.data.frame(read_file("protected_data/faulkner_absalom_parentheses.txt"))
colnames(parentheses_file) <- "text"

parentheses_clean <- parentheses_file %>%
  mutate (text = gsub("[\t\n\r]", "", text, perl = TRUE)) 
 
#This string is used for the Python script
absalom_string <- parentheses_clean$text
```

```{python nested_sentences}
#This code snippet converts all of the nested parentheses into a list object. Full credit goes to: https://stackoverflow.com/users/68063/gareth-rees

def parenthetic_contents(string):
    """Generate parenthesized contents in string as pairs (level, contents)."""
    stack = []
    for i, c in enumerate(string):
        if c == '(':
            stack.append(i)
        elif c == ')' and stack:
            start = stack.pop()
            yield (len(stack)+1, string[start + 1: i],start, start + len(string[start + 1: i]))

py_nested_sentences = list(parenthetic_contents(r.absalom_string))
```

```{r get_python_value}
#Create Parenthetical_df for processing this. 
#Convert the Python list object into a 2 vector data frame. This gave me no end of trouble because the vector kept turning into a list. Could not find a tidy function for this.

nested_sentence <- py$py_nested_sentences

parenthetical_df <-
  as.data.frame(matrix(
    unlist(nested_sentence),
    nrow = length(nested_sentence),
    byrow = TRUE
  ))

colnames(parenthetical_df) <- c("depth", "sentence", "start", "end")
```

```{r create_parenthetical_levels}
#Clean up table
parenthetical_level <- parenthetical_df %>%
  select(-sentence) %>%
  mutate(across(.cols = everything(),  ~ as.numeric(.)))

```

```{r all_depths}
#get the values from parenthetical level_one. 

parenthetical_level_one <- parenthetical_level %>%
  filter(depth == 1)

#Get all start end blocks "outside" level one (start-1, end+1)
#Pivot into one column so start and end alternate
# 1 and total absalom lenght
#arrange by start

parenthetical_main_text <- parenthetical_level_one %>%
  mutate (depth = 0) %>%
  mutate (start = start - 1) %>%
  mutate (end = end + 1) %>%
  pivot_longer(2:3, values_to = "start") %>%
  select (-name) %>%
  add_row(depth = 0, start = 1) %>%
  add_row(depth = 0, start = str_length(absalom_string)) %>%
  arrange(start)
```

```{r combine_parenthetical_levels}
#pivot start end long, and drop start_stop values
parenthetical_long <- parenthetical_level %>%
  pivot_longer(2:3, names_to = "stops", values_to = "start") %>%
  select(-stops)

#add main text parenthetical to long parenthesis table
parenthetical_all <- parenthetical_long %>%
  bind_rows(parenthetical_main_text) 
```

```{r stagger_parenthetical_levels}
#Number of empty rows to insert
N = 1

#Every N rows after which empty rows should be inserted
after_rows = 2

parenthetical_all_staggered <- do.call(
  rbind, 
        lapply(
                split(parenthetical_all,  
                ceiling(1:NROW(parenthetical_all)/after_rows)),
                function(a) rbind(a, replace(a[1:N,], TRUE, NA))
               )
        )

parenthetical_all_staggered <- parenthetical_all_staggered %>%
  remove_rownames() %>%
  mutate(depth = ifelse(is.na(depth), lag(depth), depth))
```

```{r chapter_index}
#Create a chapter index of when each character occurs. This is useful for keeping track of the data when plotting it.

##Unnest everything by chapter and add row numbers for chapters. Then create a helper column to get the string lenght of each segment. Then get the cummulative sum with a lag of 1.

chapter_index <- parentheses_clean %>%
  unnest_regex(chapter_text, text , "Chapter \\d") %>%
  mutate(chapter = row_number()) %>%
  mutate(length = str_length(chapter_text)) %>% 
  mutate(chapter_marker = cumsum(lag(length,default=1)))
```

```{r longest_break}
#find the opening phrases of the longest break for annotation. 

parenthetical_longest_break <- parenthetical_df %>%
  mutate(across(c(1, 3:4),  ~ as.numeric(.))) %>%
  mutate (gap = end - start) %>%
  arrange(depth, desc(gap)) %>%
  group_by(depth) %>%
  top_n(gap, n= 1) %>%
  mutate (sentence = paste("(", sentence, ")", sep = "")) %>%
  mutate (sentence = ifelse(
    str_count(sentence, "\\s") > 1,
    paste(word(sentence, 1, 7), "...", sep = ""),
    sentence
  ))
```

```{r parenthetical_length}
parenthetical_sentences <- parenthetical_df %>%
  select(-start,-end) %>%
  filter(depth == 1) %>%
  mutate(word_count = str_count(sentence, "\\S+")) %>% 
  group_by(depth) %>% 
  summarize(total = sum(word_count)) %>%
  mutate(percent = total / absalom_length)
```

The typography of a novel is rarely remarkable, but here, too, *Absalom, Absalom!* distinguishes itself. Faulkner uses what can only be considered a life time's supply of parentheses -- `r str_count(absalom_string, "[()]")` total or `r str_count(absalom_string, "[()]")/2` pairs.[^8] These parenthetical statements are used to elucidate, complicate, contradict, and trouble, the "main narrative." They are so frequent and lengthy that a full `r percent(parenthetical_sentences$percent[1])` of the text is enframed by parentheses. Faulkner adds a layer of complexity to this by "nesting" parentheses -- putting parentheses inside other parentheses. He does this throughout the text, but most intensely in chapter 6, where the majority of the chapter is actually a parenthetical. Both the frequency and the nesting have been represented visually through the figure below.

[^8]: This number exceeds that of the "fallible" manual count done by Fred Randel quite a bit. He arrived at 594 pairs, but does not have the benefit computational exactitude [-@randelParenthesesFaulknerAbsalom1971 471].

```{r plot_parentheses, out.width="768px",  out.height="250px", fig.cap= "Figure 4: Snapshot of all parentheses in <i>Absalom, Absalom!</i> ordered by their level of nesting."}

plot_parentheses <- plot_ly(
  parenthetical_all_staggered,
  x = ~ start,
  y =  ~ depth,
  name  = ~ depth,
  type = 'scatter',
  mode = 'lines+markers',
  marker = list(size = 4),
  line = list(width = 1, dash = "dot"),
    width = 768,
    height = 250
)

plot_parentheses <- plot_parentheses %>%
  layout(
    title = list(text = "Parentheses Nesting Levels in <i>Absalom, Absalom!</i>"),
    xaxis = list(
      gridwidth = 1,
      zeroline = FALSE,
      showline = FALSE,
      tickmode = 'array',
      tickvals = chapter_index$chapter_marker,
      ticktext = chapter_index$chapter,
      title = list(text = "Chapter")
    ),
    yaxis = list(
      title = list(text = "Nesting Level"),
      showgrid = FALSE,
      zeroline = FALSE,
      showline = FALSE,
      linewidth = 0,
      tickmode = 'array',
      tickvals = c(0, 1, 2, 3, 4),
      ticktext = c("Main Text", "1: ()", "2: (())", "3: ((()))", "4: (((())))")
    ),
    colorway = faulkner_colorway_bw,
    font = plot_font,
    margin = m,
    autosize = FALSE
  )

plot_parentheses <- plot_parentheses %>%
  add_annotations(
    x = parenthetical_longest_break$start,
    y = parenthetical_longest_break$depth,
    text = parenthetical_longest_break$sentence,
    xref = "x",
    yref = "y",
    yshift = 15,
    xanchor = "left",
    showarrow = FALSE,
    font = plot_font
)

plot_parentheses <-  plot_parentheses %>% 
  config(
    toImageButtonOptions = list(
      format = "svg",
      width = 768,
      height = 250
    )
  )

plot_parentheses

```

The figure shows several salient features of the text purely based on its typography. First, at the bottom is the "main text" -- the text that is not parenthetical. To some extent this appears indistinguishable from the first nesting level: (). In part, this is simply the consequence of trying to plot too many data points into a small space, but it also demonstrates the ubiquity of parenthetical phrases. With some regularity, this first level nesting is further enframed with second level nesting: (()). The most intense usage of nesting is in chapter 6. In the first couple of pages, the narrative jumps to level one nesting, then second level, and even third ((())), until finally at the apex of the chart, Quentin is nested four levels deep, or more colloquially, he is an aside to an aside to an aside to an aside. It is likely a happy accident that Quentin's name is the only one that is so deeply nested, it is definitely unreasonable to expect any reader to discover this Easter egg without considerable effort. To be sure, it is possible that Faulkner himself lost track of the parentheses as the number of parentheses in this chapter is uneven. Formally, there should be an extra closing parenthesis at the end of chapter 6. Be that as it may, to find Quentin dead-center in the text at the deepest parenthetical level reinforces how much *Absalom, Absalom!* is as much about his confrontation with the past, as it is about the troubled history of the Sutpens. In many respects that makes him the main character, or as Alex Vernon points out, "Every other voice in the text comes to us through Quentin" [@vernonNarrativeMiscegenationAbsalom2001 164].

Along with Quentin's deep nesting, the liberal use of parentheticals also speaks to the text's relationship with storytelling and history. Just as the narrators mix and overlap, so the story constantly branches off from the main narrative. These branches continuously draw different information into the story. It is a typographical indicator that no story can ever truly be told one way, but must be told several ways simultaneously.

```{r unique_words}
unique_words <- absalom_df %>%
  unnest_tokens(word, text) %>%
  distinct(word) %>%
  filter (word == str_match(word, "[a-z]+"))

word_list <- as.data.frame(read_file("data/words_alpha.txt"))
colnames(word_list) <- "word"

word_list <- word_list %>%
  unnest_tokens(word, word)

neologisms_export <- unique_words %>%
  anti_join(word_list)
write_csv(neologisms_export, "data/neologisms_export.csv")

neologisms_import <- read_csv("data/neologisms_import.csv")
neologisms_count <- nrow(neologisms_import)

unique_words_count <- unique_words %>%
  nrow()
```

This sense of simultaneity is reinforced through the text's repetitiveness. Words, phrases, scenes, and narrators echo one another throughout the text and evoke the sense that one is rereading a particular passage several times [@rio-jelliffeAbsalomAbsalomSelfReflexive1981d 76]. This effect can be observed in the data about the text as well. For all this apparent lexical originality, statistically speaking, *Absalom, Absalom!* does not stand out for its word use. There are `r unique_words_count` unique words, which is actually fewer unique words than expected for a text of its length, making it slightly more repetitious [@chacomaHeapsLawHeaps 4; @bernhardssonMetaBookSizedependent2009 14]. [^9]

[^9]: Heaps' law predicts that that the number of unique words in a text is approximately the square root of the total number of words in a text. $V_R(n)= Kn^\beta$ where $V_R(n)$ is the number of distinct words in corpus $n$. The special parameters $10<K<100$ and $\beta \approx .5$ vary from corpus to corpus.

```{r narrative_language, eval=FALSE}
# Grab all ngrams between 4 and 16 words long. These numbers were arrived at using manual fine tuning. Filter out duplicate phrases by count>1. Add the ngram length.
absalom_ngram <- absalom_df %>%
  unnest_ngrams(word, text, n_min = 4, n = 16) %>%
  count(word) %>%
  filter(n > 1) %>%
  mutate (ngram_length = str_count(word, "\\S+"))

deduped <-
  #the map function iterates a function across the entire vector. In this case, search for each string in each string and keep track of the count.
  map_df(
    absalom_ngram$word,
    ~ str_count(absalom_ngram$word, .x) %>%
      sum  %>%
      set_names(.x) %>%
      enframe(name = "word", value = "count")
  ) %>%
  arrange(desc(count))

#filter out any duplicates.
deduped_short <- deduped %>%
  filter (count == 1)

#filter out the duplicated ngrams
ngrams_filtered <- absalom_ngram %>%
  inner_join(deduped_short)

#recount how often each individual ngram appears in the text. There is a discrepancy between the initial ngram count and their actual appearance in the text. Not sure why, but this solves that issue.
ngrams_text_count <- ngrams_filtered %>%
  mutate(text_count =  str_count(absalom_df$text, word))
```

```{r manual_store_ngrams}

#write_csv(ngrams_text_count, "data/ngrams_text_count.csv")

ngrams_text_count <- read_csv("data/ngrams_text_count.csv")

```

```{r n_grams_text count}
#Calculate text lengths.
absalom_length <- absalom_df %>%
  unnest_tokens(word, text) %>%
  nrow()

ngrams_word_count <- ngrams_text_count %>%
  mutate(total_count = ngram_length * text_count) %>%              
  group_by(ngram_length) %>%
  summarise(total = sum(total_count)) %>%
  arrange(desc(ngram_length)) %>%
  mutate(cum_total = cumsum(total)) %>%
  mutate(percent = total / absalom_length) %>%
  mutate(cum_percent = cum_total / absalom_length)

ngrams_word_count <- ngrams_word_count %>%
  mutate(title = "absalom")   
```

```{r absalom_character_ngrams}
character_names <- absalom_events_short %>% 
                    unnest_tokens(word, CharacterName) %>% 
                    distinct(word)


major_character_words <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  filter(CharacterName %in% c("Charles Bon",
                              "Judith Sutpen",
                              "Thomas Sutpen",
                              "Henry Sutpen")) %>%
  group_by(CharacterName) %>% 
  unnest_tokens(word, event_sentence) %>% 
  anti_join(stop_words) %>% 
  anti_join(character_names) %>% 
  count(word) %>% 
  top_n(5) %>% 
  arrange(CharacterName, desc(n))
  
major_character_ngrams <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  filter(CharacterName %in% c("Charles Bon",
                              "Judith Sutpen",
                              "Thomas Sutpen",
                              "Henry Sutpen")) %>%
  group_by(CharacterName) %>% 
  unnest_ngrams(word, event_sentence, n=4) %>% 
  count(word) %>% 
  top_n(5) %>% 
  arrange(CharacterName, desc(n)) 


major_character_ngrams_ind <- major_character_ngrams %>% 
  ungroup() %>% 
  add_count(word, name = "repeated_ngram") %>% 
  filter(repeated_ngram<2)

major_character_ngrams_shared <- major_character_ngrams %>% 
  ungroup() %>% 
  add_count(word, name = "repeated_ngram") %>% 
  filter(repeated_ngram>1)

major_narrators_ngrams <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  filter(CharacterName %in% c("Quentin Compson",
                              "Mr. Compson",
                              "McCannon, Shreve",
                              "Rosa Coldfield")) %>%
  group_by(CharacterName) %>% 
  unnest_ngrams(word, event_sentence, n=4) %>% 
  count(word) %>% 
  top_n(5) %>% 
  arrange(CharacterName, desc(n))

```

This repetition is more pronounced when looking at word clusters or n-grams, where *n* indicates the number of words in the phrase. Thus, an n-gram can be two words: Charles Bon, or it can be much longer. The longest n-gram in the novel is the sixteen-word sentence, "Then I could give you a decent stall in the stable' and turned and went out." It is repeated verbatim twice (151, 229). Meanwhile, the four-word phrase, "the two of them" is repeated twenty-nine times. In fact, roughly `r percent(ngrams_word_count$percent[13])` of the texts consists of repeating phrases of four to sixteen words. This number is a bit misleading though, as it follows a more general pattern of n-gram distribution in English language texts [@haExtendingZipfLaw2009 106].[^10] It also is not clear what repetitions are unique to the text, and which are simply stock phrases in English.

[^10]: It appears that n-grams also follow Zipf's law. That is, the word frequency of words is proportional to their rank order. Thus, the second most frequent word will be half as frequent as the the first most frequent, the third most frequent will be one third as frequent and so forth.

In order to highlight the n-grams that are directly relevant to *Absalom, Absalom!*, the n-grams can be cross-matched with the character around which they occur. This analysis reveals that some characters share common phrases, Charles, Henry, Judith, and Sutpen all appear in events with the phrase "the two of them." The Sutpen men all share the common n-gram "he would have to." Meanwhile, characters also have signature repetitions. For obvious reasons, "to the West Indies" is repeated around Sutpen eight times. More tellingly, the most common unique phrase for Charles Bon is "he was up to something," which is repeated seven times. The sustained accretion of certain lexical cues surrounding specific characters, models them throughout the text.

Moreover, the constant repetitions also draw attention to the narrativity of narration. Since signature phrases, scenes, and whole lines of inquiry repeat throughout the text, the meaning of each instance changes with each iteration. The same language is constantly cast in a new light, demonstrating that what is being said cannot be separated from how it is being said. It is the linguistic corollary to the repetitions in the plot structure that continuously ask the reader to reevaluate the grounding assumptions of their knowledge.

## Conclusion

The preceding vignettes cover only a fraction of the type of analysis of *Absalom, Absalom!* that is possible with the *DY* data. In some sense, this is property inherent to all databases: they make relationships between entities explicit that would otherwise be hard to detect. In another sense, the multiplicity of analyses is also idiosyncratic to *Absalom, Absalom!*, a text that categorically resists closure and requires constantly new interpretations. Nevertheless, it is possible to draw a large circle around its major themes and show how these cluster together. The characters, events, and language all try to reconcile the South's racial past with its present only to suggest that this reconciliation is not possible. There will always be a disjunction between that which happened and that which is now happening.

A common way to deal with this disjunction is to talk over it by telling a story, usually one with a happy ending. The storytellings in *Absalom, Absalom!* underscore that the telling cannot be separated from the tale and the teller. How a narrator speaks of characters is dependent on what character the narrator identifies with, the understanding of the past depends on the time the present is in, and the language spoken with depends on the listener spoken to. In the end, there is no conclusion to the conversation about race in America, only repetition. The conversation appears to be the point, though. By never allowing the text to truly ever close, Faulkner forces readers to critically re-examine their own relationships with the past, and, in turn, their relationships with themselves. It is only through this examination of the self that an understanding of the other becomes possible.

## References
