---
title: "locations"
author: "Johannes Burgers"
date: "7/8/2021"
output: html_document
---
## Part 2: Locations

### "the *Be Sutpen's Hundred* like the oldentime *Be Light*"

```{r sutpen_location}
#Percentage by location and then the Sutpen place. All Sutpen place locations are consolidated into one.
locations <- absalom_events_short %>%
  mutate(new_location = ifelse(
    str_detect(LocationTitle, "Sutpen") == TRUE,
    "Sutpen Place",
    LocationTitle
  )) %>%
  distinct(EventID, .keep_all = TRUE) %>% 
  group_by(new_location) %>%
  summarise(total_words = sum(event_words)) %>%
  mutate(percent = total_words / sum(total_words)) %>%
  filter(new_location == "Sutpen Place")

```

```{r outofYok}
out_of_yok_locations <- absalom_events_short %>%
  mutate(outyok = ifelse(X > 2300, "OutofYok", "InYok"
    )) %>%
  distinct(EventID, .keep_all = TRUE) %>% 
  group_by(outyok) %>%
  summarise(total_words = sum(event_words)) %>%
  mutate(percent = total_words / sum(total_words)) 
```

```{r sutpen_places}
#Sutpen place by chapter
sutpen_place <- absalom_events_short %>%
  mutate(new_location = ifelse(
    str_detect(LocationTitle, "Sutpen") == TRUE,
    "Sutpen Place",
    LocationTitle
  )) %>%
  filter(new_location == "Sutpen Place") %>%
  distinct(EventID, .keep_all = TRUE) %>%
  group_by(chapter) %>%
  summarise(total_words = sum(event_words)) %>%
  mutate(percent = total_words / sum(total_words)) 
```

```{r location_order}
#This was a helper table to sketch out a trajectory.
location_order <- absalom_events_short %>%
  filter(PresentMentioned == "Present") %>%
  filter(CharacterName == "Thomas Sutpen") %>%
  mutate(new_location = ifelse(
    str_detect(LocationTitle, "Sutpen") == TRUE,
    "Sutpen Place",
    LocationTitle
  )) %>%
  filter(new_location != "Sutpen Place") %>%
  relocate(new_location, .before = 1)
```

Sutpen's rise, fall, and the long shadow he casts, parallel that of slave capitalism. This economic system relied on the ceaseless expropriation of land and the routine torture of a racialized labor force to maximize agricultural production [@baptistHalfHasNever2016 113; @schermerhornUnrequitedToilHistory2018 34]. In the American South, the crop of choice was cotton. Though it has traditionally been portrayed as a "peculiar institution" in a backward region, new scholarship has shown that slave capitalism was the commercial engine that shaped the modern transatlantic world [@beckertSlaveryCapitalismNew2016 6]. Faulkner's own "peculiar" history of Sutpen's rise to power through this system, reveals slavery was neither confined to the carceral space of the plantation or the racial apartheid regime of the South, but extended along global networks of trade. True, Sutpen's Hundred and its environs are the central locus for a plurality of the action (`r percent(locations$percent[1])`), but just over a third of the events take place outside of the county (`r percent(out_of_yok_locations$percent[2])`). One way of representing this narrative spatially is to view the plantation as a key node of a network that stretches across space and time. The resulting constellation underscore how the Plantation cannot be imagined without the slave capitalism system that engendered it.

Narratively, the text slowly expands the reach of the network through the various places the novel travels. Sutpen shows up in town one day and after buying his land, he is suspected of leaving for the Mississippi River to the "steamboats full of gamblers and cotton- and slave dealers to replenish his cache" (26). Quite some time elapses before he leaves the county again, and travels to New Orleans to investigate Bon (82). Shortly afterwards, the war breaks out, and he reaches as far east as Virginia (100). Starting in chapter 7, the narrative travels to Virginia again, when Shreve and Quentin start reconstructing Sutpen's early years in their monastic Harvard dorm (179). From thereon Sutpen travels to Haiti (193), only to come full circle and return to Yoknapatawpha. Even a fairly crude understanding of geography notes the expanding distances Sutpen's narrative travels: the Mississippi River ($\approx$ 80 mi.), New Orleans ($\approx$ 300 mi.), Virginia ($\approx$ 750 mi.), Haiti ($\approx$ 1550 mi.). The Sutpen family history is ostensibly a local curiosity, but its full story cannot be told without connecting it to a global network of places. This network could equally well stretch to places not directly part of Sutpen's story but mentioned nonetheless. This includes Harvard ($\approx$ 1,000 mi.), Shreve's birthplace of Edmonton, Alberta ($\approx$ 1,700 mi.), the French Architect's home of Martinique ($\approx$ 2,300 mi.), the Haitian Sugar planter owner's birth country, France ($\approx$ 4,500 mi.), and, even back to the "inscrutable dark continent" of Africa (202) ($\approx$ 6,000 mi.).

As *DY* is fundamentally a mapping project, it is possible to take this information and project it onto a digital interpretation of Yoknapatawpha. The custom map below overlays three sets of data: all event locations, locations with a global connection, and locations where Sutpen is present. The primary layer draws a circle around each location proportional to how much time the text spends there. The second layer adds those locations that have a global connection to the text. No events take place in Edmonton, Martinique, France, or West Africa, but the text mentions them. Finally, the location circles have been overlaid with "o-times" symbols ( $\otimes$ ) to indicate the locations where Sutpen appears in an event. When the underlying circle and the symbol coincide exactly Sutpen is present at all of the events at that location.

```{r locations_map}
location_frequency <- absalom_events_short %>%
  distinct(EventID, .keep_all = TRUE) %>%
  select(LocationTitle, X, new_y, event_words) %>%
  group_by(LocationTitle) %>%
  mutate(location_sum = sum(event_words)) %>%
  distinct(LocationTitle, .keep_all = TRUE) %>%
  mutate(name = "All Locations") %>%
  select(-event_words) %>%
  ungroup() %>%
  add_row(
    LocationTitle = "Martinique",
    X = 2350,
    new_y = -1353,
    location_sum = 10,
    name = "Global Connection"
  ) %>%
  add_row(
    LocationTitle = "Edmonton, Alberta",
    X = 2310,
    new_y = -1280,
    location_sum = 10,
    name = "Global Connection"
  ) %>%
  add_row(
    LocationTitle = "France",
    X = 2475,
    new_y = -1300,
    location_sum = 10,
    name = "Global Connection"
  ) %>%
  add_row(
    LocationTitle = "Africa",
    X = 2460,
    new_y = -1353,
    location_sum = 10,
    name = "Global Connection"
  ) %>%
  mutate(X = X - 632) 
```

```{r hybrid_locations}
sutpen_locations <- absalom_events_short %>%
  filter(CharacterName == "Thomas Sutpen") %>%
  filter(PresentMentioned == "Present") %>%
  distinct(EventID, .keep_all = TRUE) %>%
  select(LocationTitle, X, new_y, event_words) %>%
  group_by(LocationTitle) %>%
  mutate(location_sum = sum(event_words)) %>%
  distinct(LocationTitle, .keep_all = TRUE) %>%
  mutate(name = "Thomas Sutpen") %>%
  mutate(X = X - 632) %>%
  select(-event_words, -X, -new_y)

hybrid_locations <- location_frequency %>%
  full_join(sutpen_locations,
            by = "LocationTitle",
            suffix = c("_all", "_sut")) %>%
  mutate(name_sut = "Thomas Sutpen") 
```

```{r map_annotations}
#write_csv(hybrid_locations, "hybrid_locations_export.csv")
#Manual clean for locations

map_annotation <- read_csv("hybrid_locations.csv")
```

```{r plot_locations_map, out.width="768px",  out.height="605px", fig.cap="Figure 2: All locations in <i>Absalom, Absalom!</i>. The map underscores the global scale of the novel."}


image_file <-
  "https://raw.githubusercontent.com/joostburgers/absalom_sentiment_analysis/master/images/dy_base_layer_4_2.png"

plot_locations <-
  plot_ly(
    hybrid_locations,
    x = ~ X,
    y = ~ new_y,
    name =  ~ name_all,
    #colors= ~name_all,
    type = 'scatter',
    mode = 'markers',
    size = ~ location_sum_all,
    # sizes = c(5, max(hybrid_locations$location_sum)*.003),
    hoverinfo = 'text',
    text = ~ paste('Total event length:', location_sum_all, '<br>',  
                   LocationTitle),
    marker = list(
      sizemode = 'diameter',
      opacity = .8,
      line = list(width = 1, color = 'white'),
      symbol = c("circle")
    )
  )

plot_locations <- plot_locations %>%
  add_trace(
    size = ~ location_sum_sut,
    name = ~ name_sut,
    marker = list(
      sizemode = 'diameter',
      opacity = .8,
      line = list(width = 1, color = 'black'),
      symbol = "circle-x-open",
      color = 'black'
    )
  )

plot_locations <- plot_locations %>%
  layout(
    title = "Locations in <i>Absalom, Absalom!</i>",
    xaxis = list(
      title = "",
      visible = FALSE,
      fixedrange = TRUE,
      range = c(0, 2008)
    ),
    yaxis = list(
      title = "",
      visible = FALSE,
      fixedrange = TRUE,
      range = c(-1660, 0)
    ),
    margin = m,
    font = plot_font,
    showlegend = TRUE,
    legend = list(
      title = list(text = "<b>Locations</b>"),
      itemsizing = 'constant'
      
    ),
    autosize = F,
    width = 768,
    height = 605,
    images = list(
      list(
        source =  image_file,
        xref = "x",
        yref = "y",
        x = 0,
        y = -40,
        sizex = 2008,
        sizey = 1660,
        sizing = "stretch",
        opacity = 0.8,
        layer = "below"
      )
    ),
    colorway = c("#C0C0C0", "#101010")
  )

plot_locations <- plot_locations %>%
  add_annotations(
    x = map_annotation$X,
    y = map_annotation$new_y,
    xref = 'x',
    yref = 'y',
    showarrow = TRUE,
    showarrow = TRUE,
    arrowhead = 3,
    arrowwidth = 1,
    arrowsize = .5,
    xanchor = 'left',
    ax = map_annotation$xoffset,
    ay = map_annotation$yoffset,
    text = map_annotation$LocationTitle,
    font = list(size = 12),
    align = 'left'
    
  )

plot_locations <-  plot_locations %>% 
  config(
    toImageButtonOptions = list(
      format = "png",
      width = 768,
      height = 605
    )
  )



plot_locations



```

Reading the map from left to right, several features stand out. The Sutpen Plantation is the most prominent location, and, in fact, many of the smaller circles in its immediate vicinity are places on the vast property. Within the county, most of the locations are confined to the areas between the plantation and the town, with a smattering of events in the direct north of the county when Sutpen is searching for the French Architect. Taking a broader view, the action travels to quite a number of places outside of Yoknapatawpha, notably New Orleans, Tidewater, the South during the Civil War, and Haiti. Sutpen is ubiquitous across these locations and there are few places where he is the most present character in an event, or indeed, the only present character in an event.

Beyond these locations, the global connections enmesh the text in the larger transatlantic circuit. The extensive distances between these locations suggests that slave capitalism is not only maintained through intimate social relations, but also reticulated across a network of interlocking spaces. The various characters use these spaces to continuously displace the evils of slavery onto a place "somewhere else," thereby keeping their own moral complicity at a distance. Rosa places the enslavement in Yoknapatawpha on a continuum with an "old quiet country like Virginia or Carolina," as if its continued practice were a natural evolution (11). Charles Bon's relationship with an "octoroon" mistress is possible because New Orleans is a sinful place, and the home of a thousand men "without honor or principle" who produced these women and ensnare them into concubinage (91). The town fears Sutpen's "wild stock" of African captives who speak with a foreign tongue will mix with their "tame" enslaved peoples (48). Meanwhile, Quentin's grandfather and, in turn, his father, describe Haiti as, "the halfway point between what we call the jungle and what we call civilization" (202). Rather than Sutpen's Hundred, the "little island" becomes the true "theatre for violence and injustice" (202). The original sin of slavery is again displaced from the town, and the South more generally. This time onto an Edenic and infernal island in the Caribbean.

The displacement of slavery onto Haiti is striking. Haiti became the first Black republic in the Western Hemisphere in 1804, and thus abolished slavery several years before Sutpen was even born. This historical discrepancy has spawned a substantial critical debate [See for example: @goddenAbsalomAbsalomHaiti1994; @matthewsRecallingWestIndies2004; @stanchichHiddenCaribbeanOther1996]. Most recently, Wanda Raiford has made the provocative argument that the elision of the historical Haiti and its unqualified acceptance by every listener in the text, also implicates the reader in the erasure of US imperial history and Black agency [-@RN27 103]. Extending this logic, Haiti becomes the primordial site onto which to project the evils of slavery. In doing so, General Compson and indeed the entire Compson line, deflect any historical guilt for their own violent subjugation of other human beings. If anything, the "wild" form of slavery in the non-existent version of Haiti, sanctions the "tame" enslavement in the "civilized" US South.

As the constellation of locations suggests, slavery capitalism does not stop neatly at the Mason-Dixon line, and also undergirds Quentin's attendance at Harvard. Despite its role in the abolitionist movement, Harvard directly and indirectly profited from slavery [@RN40 226]. Moreover, Quentin is following in the footsteps of other sons of slave owners and, later, former slave owners who went to Harvard or other elite Northern institutions to be conferred a degree and respectability [-@RN40 238]. Though not explicitly stated in *Absalom, Absalom!*, in *The Sound and the Fury* it is clear that the tract of land Quentin's father sells to send his son to school belonged to the Compson plantation. The residual wealth from a system of subjugation that officially ended fifty years previous still helps pay for his tuition.

Only Shreve appears untouched. As a Canadian outsider and someone unfamiliar with the South, he has no historical link with slavery. Yet, he and Quentin are still connected by the same River, the "Continental Trough" and "geologic umbilical" that runs from Mississippi to Edmonton (208). The waterways that bring together the two Harvard classmen were also used to ferry enslaved people, the commodities they produced, and the ideas that perpetuated these social relations. Shreve has inherited the slave capital's perverse racial logic through these various arteries. While Quentin is famously ambivalent about his relationship with the South and its troubled past, Shreve's vision of the region and the future it portends is one of white racial suicide. He prophesies that "the Jim Bonds are going to conquer the western hemisphere" and fears that in "a few thousand years, I who regard you will also have sprung from the loins of African kings" (302). It is unclear how much of this is Shreve's trademark flippancy, but even in his glib sermonizing about the fate of the Western world, he reveals profound anxieties over the disappearance of "pure" White people. He claims that the future Jim Bonds might "bleach out" as they reach the poles, but it will "still be Jim Bond" suggesting that Black heritage represents an inferiority beyond skin color (303). Shreve's own obsessions with race suggests that the legacy of enslavement cannot be told through one individual family, region, or, even country, but is, in fact, the origin story of the Western world itself.


